BayesPy: Variational Bayesian Inference in Python

### 2015-03-13 response.tex

% \end{itemize}

\subsection{Reviewer 1}

\begin{itemize}

\item \underline{supported models}
  \\
  The revised version makes it more clear that BayesPy supports conjugate
  exponential family models and does not currently support non-conjugate models.
  An example of a non-supported model is given in section 4.  Examples of
  supported popular models are listed in section 2.  Section 3 also provides an
  example implementation of a Gaussian mixture model, which is much more
  interesting model than the very simple model in the previous version of the
  paper.  Because of the page limit, more examples are provided in the online
  documentation (bayespy.org).  Section 4 also discusses Gaussian mixture and
  PCA models, which are given implementations in the supplementary material.
  The results in the section also provide some rough idea about computational
  costs and how large models can be constructed without using stochastic
  variational inference.

@@

\item \underline{speed comparison}
  \\
  The revised version contains a more detailed comparison with Infer.NET.  The
  other relevant active packages are basically MCMC packages.  Because it is not
  obvious how to weigh different aspects of MCMC and VB methods in general, a
  comparison with the MCMC packages is out of the scope of this paper.  However,
  because Infer.NET seems to offer practically the same functionality, section 4
  provides a detailed comparison between BayesPy and Infer.NET, including a
  speed comparison.

\item \underline{scheduling}
  \\
  The updates (and all computations) are performed sequentially, that is,
  variables are updated in turns.  This is now mentioned in Section 3.  Parallel
  computing is not supported.

\item \underline{syntax}
  \\
  The API in the online documentation provides details about the syntax for
  nodes and methods.  For instance, more details about the syntax of the gating
  node has been added in the documentation.  However, it is typically better to
  use Mixture node directly to achieve mixture distributions, as shown in the
  example in section 3.

### 2015-03-16 bayespy.tex
Revised submission

Several other projects have similar goals for making Bayesian inference easier
and faster to apply.  VB inference is available in Bayes Blocks
\citep{Raiko:2007}, VIBES \citep{Bishop:2002} and Infer.NET \citep{Infer.NET}.
Bayes Blocks is an open-source C++/Python package but limited to scalar Gaussian
nodes and a few deterministic functions, thus making it very limited.  VIBES is
an old software package for Java, released under the revised BSD license, but it
is no longer actively developed.  VIBES has been replaced by Infer.NET, which
%supports a wide range of posterior approximations.  However, Infer.NET
is partly closed source and licensed for non-commercial use only.  Instead of VB
inference, mainly MCMC is supported by other projects such as PyMC \citep{PyMC},
OpenBUGS \citep{OpenBUGS}, Dimple \citep{Dimple} and Stan \citep{Stan}.  Thus,
there is a need for an open source and maintained VB software package.

@@

This section provides a brief comparison with Infer.NET because it is another
active project implementing the variational message passing algorithm, whereas
the other related active projects implement MCMC.  The main advantages of
Infer.NET over BayesPy are its support for non-conjugate models (e.g., logistic
regression) and other inference engines.  On the other hand, BayesPy is
open-source software and supports several advanced VB methods (e.g., stochastic
variational inference and collapsed inference).

@@

The results are summarized in Table~\ref{tab:speed}.  On the smaller datasets,
Infer.NET is slightly faster because BayesPy has some Python overhead.  However,
the total CPU time for both methods is only a few seconds.  On the larger
datasets, BayesPy is significantly faster because it uses BLAS and LAPACK
libraries for fast numerical operations.\footnote{Assuming that NumPy uses
  BLAS/LAPACK because BayesPy uses NumPy for numerical operations.}  In
addition, if using multi-threaded BLAS and LAPACK on a multi-core computer, the
difference in wall-clock time can be even larger than in CPU time.  The speed of
BayesPy for these models could potentially be improved by using collapsed or
stochastic variational inference.  The accuracy could be improved by using
Gaussian-Wishart nodes which do not factorize with respect to the Gaussian and
Wishart variables.

>>>>

Several other projects have similar goals for making Bayesian inference easier
and faster to apply.  VB inference is available in Bayes Blocks
\citep{Raiko:2007}, VIBES \citep{Bishop:2002} and Infer.NET \citep{Infer.NET}.
Bayes Blocks is an open-source C++/Python package but limited to scalar Gaussian
nodes and a few deterministic functions, thus making it very limited.  VIBES is
an old software package for Java, released under the revised BSD license, but it
is no longer actively developed.  VIBES has been replaced by Infer.NET, which
%supports a wide range of posterior approximations.  However, Infer.NET
is partly closed source and licensed for non-commercial use only.  Instead of VB
inference, mainly MCMC is supported by other projects such as PyMC \citep{PyMC},
OpenBUGS \citep{OpenBUGS}, Dimple \citep{Dimple} and Stan \citep{Stan}.  Thus,
there is a need for an open-source and maintained VB software package.

@@

This section provides a brief comparison with Infer.NET because it is another
active project implementing the variational message passing algorithm, whereas
the other related active projects implement MCMC.  The main advantages of
Infer.NET over BayesPy are its support for non-conjugate models (e.g., logistic
regression) and other inference engines (EP and Gibbs sampling).  On the other
hand, BayesPy is open-source software and supports several advanced VB methods
(e.g., stochastic variational inference and collapsed inference).

@@

The results are summarized in Table~\ref{tab:speed}.  On the smaller datasets,
Infer.NET is slightly faster because BayesPy has some Python overhead.  However,
the total CPU time for both methods is only a few seconds.  On the larger
datasets, BayesPy is significantly faster because it uses BLAS and LAPACK
libraries for fast numerical operations.\footnote{Assuming that NumPy uses
  BLAS/LAPACK because BayesPy uses NumPy for numerical operations.}  In
addition, if using multi-threaded BLAS and LAPACK on a multi-core computer, the
difference in wall-clock time can be even larger than in CPU time.  The speed of
BayesPy for these models might be improved by using collapsed or stochastic
variational inference.  The accuracy of BayesPy could be improved by using a
Gaussian-Wishart node, which does not factorize with respect to the Gaussian and
Wishart variables.
